---
title: "Peat Depth Model: Explore and Select Data"
params:
  params_fn: 
    input: file
    label: 'Select parameters to import (.rds)'
    value: "../data/parameters.rds"
output:
    html_notebook:
    html_document:
    df_print: paged

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo= TRUE)
```

-------  

Run at `r Sys.time()` for **`r params$area_name`** (*`r params$area_abbr`*) with the following parameters: 

```{r, echo=FALSE}
#import parameters from previous runs
parameters <- readRDS(file = params$params_fn)

# if you need to change any imported parameters, do it here, e.g.: 
  #parameters$existing_parameter <- newvalue
  #parameters$new_parameter <- value

# add new parameters from yaml header


# print list of all parameters
parameters
```


-------    

## Introduction

This script allows you to:  

1. Import previously prepared data:   
    * peat observations with environmental covariate values
    * rasters of environmental covariates
    * a prediction area
2. Explore the distribution of peat depths
3. Explore the covariate space
4. Stratify and or subset peat depth observations, predictors and prediction areas



```{r}
#load packages
library(tidyverse)#
library(raster)
library(rgdal)
library(sp)
library(sf)
library(elevatr)
```


## 1. Import previously prepared data

```{r}
# import input.data
input.data.sp <- readRDS(parameters$input_data.sp.fn)
input.data <- input.data.sp %>% as_tibble()

# import elevation and slope
elevation <- raster(parameters$elev_raster_fn)
slope <- raster(parameters$slope_raster_fn)

# stack predictors
pred_stack <- stack(elevation, slope)
```


## 2. Explore the distribution of peat depths

```{r}
ggplot(input.data %>% filter(depth != 0), 
       aes(x = depth)) +
  geom_histogram(binwidth = 0.1, boundary = 0) +
  #scale_x_continuous(breaks = seq(0,20), minor_breaks = NULL) +
  labs(title = "Distribution of peat depths sampled", x = "Peat depth in metres") +
  theme_light()
```

## 3. Explore the covariate space


```{r}
ggplot(input.data, aes(x = X, y = Y)) +
  geom_point(aes(colour = depth), size = 1) +
  geom_rug(length = unit(0.02, "npc"), colour = "darkgrey") +
  scale_colour_viridis_c(direction = -1, option = "plasma", end = 0.8)+
  scale_x_continuous(breaks = seq(0,600000,20000)) +
  scale_y_continuous(breaks = seq(0,1000000, 20000)) +
  theme_light() +
  coord_equal() +
  labs(title = "Peat depth spatial distribution")
```

```{r}
p.sl.el <- ggplot(input.data, aes(x = elev, y = slope)) +
  geom_point(aes(colour = depth), size = 1)+
  geom_rug(length = unit(0.02, "npc"), colour = "darkgrey") +
  scale_colour_viridis_c(direction = -1, option = "plasma", end = 0.8)+
  coord_fixed(ratio = 15) +
  theme_light() +
  labs(title = "Peat depth: elevation by slope")

filter_depth <- 200
p.sl.el.deep <- ggplot(input.data %>% filter(depth > filter_depth), 
       aes(x = elev, y = slope)) +
  geom_point(aes(colour = depth), size = 1)+
  geom_rug(length = unit(0.02, "npc"), colour = "darkgrey") +
  scale_colour_viridis_c(direction = -1, option = "plasma", end = 0.8)+
  #coord_fixed(ratio = 15) +
  theme_light() +
  labs(title = paste0("Deeper peat (>", filter_depth, "cm): elevation by slope"))

plot(p.sl.el)
plot(p.sl.el.deep)
```


```{r}
#elev v depth
ggplot(input.data, aes(x = depth, y = elev)) +
  geom_point(aes(colour = depth), size = 1)+
  geom_rug(length = unit(0.02, "npc"), colour = "darkgrey") +
  scale_colour_viridis_c(direction = -1, option = "plasma", end = 0.8)+
  #coord_fixed(ratio = 15) +
  theme_light() +
  labs(title = "Peat depth: elevation by depth")

#slope v depth
ggplot(input.data, aes(x = depth, y = slope)) +
  geom_point(aes(colour = depth), size = 1)+
  geom_rug(length = unit(0.02, "npc"), colour = "darkgrey") +
  scale_colour_viridis_c(direction = -1, option = "plasma", end = 0.8)+
  #coord_fixed(ratio = 15) +
  theme_light() +
  labs(title = "Peat depth: slope by depth")
```





## 4. Stratify and or subset peat depth observations, predictors and prediction areas




```{r}
# export parameters
saveRDS(parameters,  paste0("../data/parameters_", 
                                      parameters$area_abbr, ".rds"))
```

