---
title: "Peat Depth Model: Import and Prepare Data"
params:
  area_name: 
    input: text
    label: 'Long name of area to be modelled'
    value: Spadeadam
  area_abbr:
    input: text
    label: 'Abbreviation of the area to be modelled to label outputs'
    value: spade
  pred_bound: 
    input: file
    label: 'Prediction boundary: select shapefile (leave as-is if selecting NP)'
    value: ../data/site_box.shp
  NP: 
    input: select
    label: 'Prediction boundary: select National Park if applicable'
    value: DARTMOOR
    choices: [NA, DARTMOOR, EXMOOR, LAKE DISTRICT, NEW FOREST, NORTH YORK MOORS, NORTHUMBERLAND, PEAK DISTRICT, SOUTH DOWNS, THE BROADS, YORKSHIRE DALES]
  raster_source:
    input: select
    label: 'Source of raster data (pick TerrainTiles to automatically download terrain)'
    value: select a local GeoTiff file
    choices: [TerrainTiles, select a local GeoTiff file]
  terrain_raster_file: 
    input: file
    label: 'Terrain raster file in GeoTiff format (if not selecting TerrainTiles)'
    value: "../data/terr_spade_12.tif"
  cell_size: 
    input: numeric
    label: "Desired raster resolution in metres:"
    value: 5
  obs_path:
    input: file
    label: 'Observations data (shapefile):'
    value: ../data/peatdepthSpadeAll.shp
  projection:
    input: text
    label: 'Enter projection proj4 string (lookup at https://spatialreference.org, default is BNG):'
    value: +proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +ellps=airy +datum=OSGB36 +units=m +no_defs 
  field_depth:
    input: text
    label: "peat depth field:"
    value: "Peat_depth"

output:
    html_notebook:
    html_document:
    df_print: paged

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

-------  

Run at `r Sys.time()` for **`r params$area_name`** (*`r params$area_abbr`*) with the following parameters: 

```{r, echo=FALSE}
# start saving project parameters
parameters <- params
parameters 
```


-------    

## Introduction

This script allows you to:  

1. Import or define your prediction area
2. Download an elevation raster for your prediction area
3. Calculate a slope raster
4. Import a shapefile of peat depths
5. Extract environmental covariates for each peat depth observation
6. Export environmental covariate rasters and a table of peat depth observations and covariate values


This script downloads terrain rasters for a given area. 

```{r}
#load packages
library(tidyverse)#
library(raster)
library(rgdal)
library(sp)
library(sf)
library(elevatr)
```


##  1. Prediction boundary

```{r}
#get boundary shapefile if one is specified
if(parameters$pred_bound != "National Park"){
site_bound <- sf::st_read(parameters$pred_bound) %>% as_Spatial()}

#get NP boundary if a no boundary shapefile specified
if(parameters$pred_bound == "National Park"){
# Import site boundary from National Park
  NP_boundaries <- readOGR(dsn = "../data/NP_boundaries/National_Parks__England____Natural_England.shp", verbose = TRUE) %>% 
    st_as_sf()
  site_bound <- NP_boundaries[which(NP_boundaries$NAME == parameters$NP),] %>% as_Spatial()
}

# re-project
site_bound <- spTransform(site_bound, CRSobj = parameters$projection)

# export
saveRDS(site_bound, "../data/site_bound.rds")

#clean up
if(exists("NP_boundaries")){rm(NP_boundaries)}
plot(site_bound)
```

## 2. Obtain elevation raster

This section 

* loads a terrain raster:  
    + downloads a terrain raster from TerrainTiles, if the option is selected, and writes that terrain raster to a local GeoTiff file (named `terr_area_abbr_zoom.tif`)  
    + alternatively loads a terrain raster from a local GeoTiff if a file path is provided.    
* resamples the raster to the desired resoution `cell_size` and reprojects it
* crops the terrain raster to the prediction area.  
* creates a slope raster and saves this to a local GeoTiff (named `slope_area_abbr_zoom.tif`)


```{r}
#if selected, get terrain tiles from TerrainTiles at stated zoom level. 
  # note that you can change the zoom level.  14 is most detail (about 3m resolution) but huge files
if(parameters$raster_source == "TerrainTiles"){
  zoom <- 12
  elevation_orig <- get_elev_raster(locations = site_bound, z = zoom)
  writeRaster(elevation_orig, paste0("../data/terr_", parameters$area_abbr, "_", zoom, ".tif"), overwrite = TRUE)
}

if(parameters$raster_source == "select a local GeoTiff file"){
  elevation_orig <- raster(parameters$terrain_raster_file)
  zoom <- 12
}
```

## 3. Reformat elevation and calculate a slope raster

```{r}
#create an empty raster of desired dimension and resolution to reproject and resample to

empty_raster <- raster(extent(site_bound), crs = parameters$projection, 
                       res = parameters$cell_size) 

#reproject to osgb and resample to desired cell size

elevation <- resample(x = elevation_orig, y = empty_raster)

#tidy up
rm(elevation_orig, empty_raster)

#create filenames and save to parameters
parameters$elev_raster_fn <- paste0("../data/elev_", 
                                    parameters$area_abbr, "_", 
                                    parameters$cell_size, "m.tif")

parameters$slope_raster_fn <- paste0("../data/slope_", 
                                     parameters$area_abbr, "_", 
                                     parameters$cell_size, "m.tif")

#export to tif file
writeRaster(elevation, filename = parameters$elev_raster_fn, overwrite = TRUE)

#calculate slope and export to tif file
slope <- terrain(elevation, opt="slope", unit="degrees", neighbors=8, 
                 filename = parameters$slope_raster_fn, 
                 overwrite = TRUE)

#slope <- raster(paste0("../data/slope_", parameters$area_abbr, "_", parameters$cell_size, "m.tif")) #use this if you've already got a slope file
#plot(slope)
```


## 4. Import a shapefile of peat depths

This section: 

* imports peat depth observations from a shapefile


```{r observations}
  #Import observations----
data.shp <- sf::st_read(parameters$obs_path)
data.shp <- sf::st_transform(x = data.shp, crs = CRS(proj4string(site_bound)))
#create datafram and spatial points data frame
data <- data.frame(data.shp)
coords <- st_coordinates(data.shp)
data.sp <- sp::SpatialPointsDataFrame(data = data, coords = coords, proj4string = CRS(proj4string(site_bound)))

#head(data.sp)
```

## 5. Extract environmental covariates for each peat depth observation


```{r}
#create dataframe with peat depth as first col
input.data <- data.frame(peat_depth = data[parameters$field_depth])
names(input.data) <- "depth"
# transform peat depth data
input.data$sqrt_depth <- sqrt(input.data$depth)

#raster::extract environmental and topographic data for each peat depth measurement
input.data$elev <- raster::extract(
  elevation, coordinates(data.sp)[,1:2])
input.data$slope <- raster::extract(slope, coordinates(data.sp)[,1:2])

#convert to spatial points data frame
coords <- st_coordinates(data.shp)
proj4 <- parameters$projection

input.data.sp <- sp::SpatialPointsDataFrame(data = input.data, coords = coords, proj4string = CRS(proj4))


spplot(input.data.sp)
```




```{r}
summary(input.data.sp)
```


## 6. Export table of peat depth observations and covariate values

```{r}
# create filename and save to parameters
parameters$input_data.sp.fn <- paste0("../data/input.data_", 
                                      parameters$area_abbr, "_", 
                                      parameters$cell_size, "m.rds")

# export observations and covariates table
saveRDS(input.data.sp, parameters$input_data.sp.fn)
```

```{r}
# export parameters
saveRDS(parameters,  paste0("../data/parameters_", 
                                      parameters$area_abbr, ".rds"))
parameters
```






