---
title: "Peat Depth Model: Model"
params:
  params_fn: 
    input: file
    label: 'Select parameters to import (.rds)'
    value: "../data/parameters.rds"
output:
    html_notebook:
    html_document:
    df_print: paged

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo= TRUE)
```

-------  

Run at `r Sys.time()` for **`r params$area_name`** (*`r params$area_abbr`*) with the following parameters: 

```{r, echo=FALSE}
#import parameters from previous runs
parameters <- readRDS(file = params$params_fn)

# if you need to change any imported parameters, do it here, e.g.: 
  #parameters$existing_parameter <- newvalue
  #parameters$new_parameter <- value

# add new parameters from yaml header

  # automatically defined session variables 
  #set current time and date to append to run log 
parameters$rundate <- paste0("run_", format(Sys.time(), format = "%Y%m%d-%H%M"))

# print list of all parameters
parameters
  
#set params for script
par.ori <- par() #save default plotting parameters

rundate <- parameters$rundate
```


-------    

## Introduction

This script allows you to:  

1. asdjklasdfjlasdfl;sd



```{r}
#load packages
library(tidyverse)#
library(raster)
library(rgdal)
library(sp)
library(sf)
library(elevatr)
library(geoR)
library(gstat)
library(stars)
```
## 1. Import previously prepared data

```{r}
# import input.data
input.data.sp <- readRDS(parameters$input_data.sp.fn)
input.data <- input.data.sp %>% as_tibble()

# # import elevation and slope
# elevation <- raster(parameters$elev_raster_fn)
# slope <- raster(parameters$slope_raster_fn)
# 
# # stack predictors
# predictors <- stack(elevation, slope)
# names(predictors) <- c("elevation", "slope")

# alternative import elevation and slope using stars
preds_st <- read_stars(c(elev = parameters$elev_raster_fn, slope = parameters$slope_raster_fn), proxy = T)
names(preds_st) <- c("elev", "slope")

st_bbox(preds_st)
st_crs(preds_st)
```

```{r}
#check projections are the same
st_crs(preds_st) == st_crs(input.data.sp)
```



```{r}
plot(preds_st[1])
plot(preds_st[2])
```

```{r}
ggplot() + geom_stars(data = preds_st, downsample = 100) +
  coord_equal() + theme_minimal()
```

```{r}
predictors <- preds_st
```


###Create prediction area and crop input data

```{r}
#select a well sampled area to predict within----
# plot(input.data.sp)
#     #small area: 
#     crop.extent <- extent(c(xmin = 25575, xmax = 260000, ymin = 75000, ymax = 80000))
  #all dartmoor

crop.extent <- st_bbox(predictors) %>% matrix(byrow = F, nrow = 2) %>% extent() #converts bbox to extent object for sp

plot(crop(input.data.sp, crop.extent))

input.data.cropped.sp <- crop(input.data.sp, crop.extent)

dat <- data.frame(input.data.cropped.sp@data,
                  input.data.cropped.sp@coords) %>%
  rename(x = X, y = Y)

input.data.cropped.map <- dat

#THIS keeps crashing: memory exhausted.  Filesize to big. 
# predictors.cropped.sp <- crop(predictors.sp, crop.extent)
# rm(predictors.sp)
# predictors.cropped.map <- data.frame(predictors.cropped.sp@data,
#                                      predictors.cropped.sp@coords)
```



```{r}
# CONVERT this to sf - to avoid converting tif to vector
# #graph it
# p.theme <- theme(legend.position = "right")
# p.elev <- ggplot(predictors.cropped.map, aes(x = x, y = y)) +
#   geom_point(aes(colour = elev)) + coord_equal() +
#   geom_text(data = input.data.cropped.map, label = "x", size = 0.8)+
#   scale_colour_distiller(palette = "Reds") +  
#   labs(#title = "Elevation", 
#        x = "", y = ""); p.elev + p.theme
# 
# p.slo <- ggplot(predictors.cropped.map, aes(x = x, y = y)) +
#   geom_point(aes(colour = slope))  + coord_equal()+ 
#   geom_text(data = input.data.cropped.map, label = "x", size = 0.8)+
#   scale_colour_distiller(palette = "Greens") +  
#   labs(#title = "Slope", 
#        x = "", y = ""); p.slo + p.theme
# # cowplot::plot_grid(p.elev + p.theme, p.asp + p.theme, p.slo + p.theme, axis = "left", nrow = 3, labels = "")
```

###Convert to inputs to gstat format 

```{r}
input.data.gs <- input.data.cropped.sp #spdf format

#remove duplicates 
input.data.gs <- remove.duplicates(obj = input.data.gs)

#If duplicate locations are an important part of data then better to use jitterDupCoords
summary(input.data.gs)
summary(input.data.cropped.sp)
```

### Exploratory linear model analysis

Using raw peat depth

```{r}
mod.lm.test <- lm(depth ~ elev + slope, data=dat)
summary(mod.lm.test)

resid <- residuals(mod.lm.test)
plot(resid)
qqnorm(resid)
qqline(resid)
dat$resid <- resid
```
not great



  still not fantastic

Using log of peat depth:...
```{r}
dat$log_depth <- log(dat$depth + 1)

mod.lm.test <- lm(log_depth ~ elev + slope, data=dat)
summary(mod.lm.test)

resid <- residuals(mod.lm.test)
plot(resid)
qqnorm(resid)
qqline(resid)
dat$resid <- resid
```
actually not the end of the world



Using square root of peat depth:...
```{r}
mod.lm.test <- lm(sqrt_depth ~ elev + slope, data=dat)
summary(mod.lm.test)

resid <- residuals(mod.lm.test)
plot(resid)
qqnorm(resid)
qqline(resid)
dat$resid <- resid
```

Using cube root of peat depth:...
```{r}
mod.lm.test <- lm(((depth)^(1/3)) ~ elev + slope, data=dat)
summary(mod.lm.test)

resid <- residuals(mod.lm.test)
plot(resid)
qqnorm(resid)
qqline(resid)
dat$resid <- resid
```
I kinda like the cube root.  Lets do it: 

```{r}
input.data.gs$cbrt_depth <- input.data.gs$depth^(1/3)
```



## 3. Fit a geostatistical model to the data to look for evidence of spatial autocorrelation after accounting for covariates


```{r}
sp_depth <- input.data.gs
```

### Plot the spatial object
In the top left plot the data have been split into 4 colours based on
quartiles (red is highest)

```{r}
spplot(sp_depth)
png(filename = paste0("../outputs/dat_overview_", 
                      parameters$area_abbr, "_", 
                      rundate, ".jpg"))
spplot(sp_depth)
dev.off()
```

### Fit a geostatistical model to all the data

see also: https://www.r-spatial.org/r/2016/02/14/gstat-variogram-fitting.html 

```{r}
sp_depth.v <- variogram(cbrt_depth ~ elev + slope, sp_depth)
sp_depth.vfit <- fit.variogram(sp_depth.v, 
                               model = vgm(psill = NA, model = "Mat", range = NA, nugget = NA), fit.method = 1)
plot(sp_depth.v, sp_depth.vfit)
summary(sp_depth.vfit)
```


```{r}
sp_depth.v <- variogram(cbrt_depth ~ elev + slope, sp_depth)
sp_depth.vfit <- fit.variogram(sp_depth.v, 
                               model = vgm(c("Exp", "Mat", "Sph", "Ste")), 
                               #fit.kappa = seq(.3,5,.01), 
                               fit.method = 1
                               )
plot(sp_depth.v, sp_depth.vfit)
sp_depth.vfit


```



```{r}
(model.geo.auto1 <- fit.variogram(sp_depth.v, 
                               model = vgm(c("Exp", "Mat", "Sph", "Ste")), fit.kappa = seq(.3,5,.01)))

(model.geo.auto2 <- fit.variogram(sp_depth.v, 
                               model = vgm(c("Exp", "Mat", "Sph", "Ste")), fit.method = 1))

(model.geo.exp <- fit.variogram(sp_depth.v, 
                               model = vgm(model = "Exp"), fit.method = 1))

(model.geo.sph <- fit.variogram(sp_depth.v, 
                               model = vgm(model = "Sph"), fit.method = 1))

(model.geo.mat <- fit.variogram(sp_depth.v, 
                               model = vgm(model = "Mat"), fit.method = 1))

```


  
```{r}
plot(sp_depth.v, model = model.geo.sph, col = "blue", main = "Spherical")
plot(sp_depth.v, model = model.geo.exp, col = "red", main = "Expontial")
plot(sp_depth.v, model = model.geo.mat, col = "green", main = "Matern")
plot(sp_depth.v, model = model.geo.auto1, col = "black", main = paste(model.geo.auto1$model[2]))
plot(sp_depth.v, model = model.geo.auto2, col = "black", main = paste(model.geo.auto2$model[2]))
```


```{r}
#chosen model
model.geo <- model.geo.auto2
summary(model.geo)
write_rds(model.geo, path = paste0("../outputs/model.geo__", 
                                   parameters$area_abbr, "_", 
                                   rundate, ".rds"))
```


```{r}
png(filename = paste0("../outputs/vgm__", 
                      parameters$area_abbr, "_",
                      rundate, ".jpg"))
plot(sp_depth.v, model = model.geo, main = rundate)
dev.off()
plot(sp_depth.v, model = model.geo, main = rundate)

```

```{r}
# export parameters
saveRDS(parameters,  paste0("../data/parameters_", 
                                      parameters$area_abbr, ".rds"))
```

