---
title: "Peat Depth Model: Predict"
params:
  params_fn: 
    input: file
    label: 'Select parameters to import (.rds)'
    value: "../data/parameters_eng.rds"
output:
    html_notebook:
    html_document:
    df_print: paged

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo= TRUE)
```

-------  

Run at `r Sys.time()` for **`r params$area_name`** (*`r params$area_abbr`*) with the following parameters: 

```{r, echo=FALSE}
#import parameters from previous runs
parameters <- readRDS(file = params$params_fn)

# if you need to change any imported parameters, do it here, e.g.: 
  #parameters$existing_parameter <- newvalue
  #parameters$new_parameter <- value

# add new parameters from yaml header

  # automatically defined session variables 
  #set current time and date to append to run log 

# print list of all parameters
parameters
  
#set params for script
par.ori <- par() #save default plotting parameters


```


-------    

## Introduction

This script predicts peat depths from models created in previous scripts.  It outputs:   

*   a raster of predicted depths
*   a raster of prediction variance



```{r}
#load packages
library(tidyverse)#
library(raster)
library(rgdal)
library(sp)
library(sf)
library(elevatr)
library(geoR)
library(gstat)
library(stars)

source("rsquared_funs.R")
source("backtransform.R")
```

```{r}
#pre-set which back-transformation to use
btf <- function(x) {
  backtransform(x = x, tf = parameters$tf)
}
```

## 1. Import previously prepared data

```{r}
#import model and data

# import model
model.geo <- readRDS("~/GitHub/Peat_Depth_Model/england/outputs/model.geo_eng_run_20200823-1202.rds")

# import observations
sp_depth <- readRDS(parameters$inputs_fn)

# columns which identify the model run
run_tag <- data.frame(area_abbr = rep(parameters$area_abbr, nrow(sp_depth)), 
                      rundate = rep(parameters$rundate, nrow(sp_depth)))
# the data used to generate the model
dat_sp <- cbind(sp_depth, 
                parameters$area_abbr, parameters$rundate)
names(dat_sp) <- c(names(sp_depth), "area_abbr", "rundate")
dat_sp$id <- seq(1,nrow(sp_depth)) #add an index
# remove duplicates 
dat_sp <- remove.duplicates(obj = dat_sp)
dat_sf <- st_as_sf(dat_sp)

# import predictors as `stars` objects
elevation_st <- stars::read_stars(parameters$elev_raster_fn)
slope_st <- read_stars(parameters$slope_raster_fn)
 
preds_st <- c(elev = elevation_st, slope = slope_st)
names(preds_st) <- c("elev", "slope")
preds_st <- st_set_crs(x = preds_st, value = st_crs(sp_depth))
#rm(elevation_st, slope_st)
```

```{r}
#check projections are the same
st_crs(preds_st) == st_crs(dat_sf)

# re-project
#dat_sp <- sp::spTransform(x = dat_sp, CRSobj = crs(sp_depth))
```

$from
[1] 1

$to
[1] 114203

$offset
[1] 82540

$delta
[1] 5


$point
[1] FALSE

$values
NULL

$from
[1] 1

$to
[1] 128824

$offset
[1] 649256

$delta
[1] -5

*reasonable tile size for kriging*
x = 10000
y = 4000
~0.3Gb per tile

```{r}
# import tiled file for testing

elev_tile <- read_stars("../data/tiles_5k/elev_eng_5m-449.tif", proxy = F)
slope_tile <- read_stars("../data/tiles_5k/slope_eng_5m-449.tif", proxy = F)

preds_tile <- c(elev_tile, slope_tile)

plot(elev_tile, reset = FALSE)
# plot(elevation_st, 
# #add = T,
#      extent = st_bbox(elev_tile))
plot(dat_sf["depth"], add = T)
```




```{r}
# tracking <- cbind(i = as.integer(), xoff = as.numeric(), yoff = as.numeric(), 
#                   Gb = as.character(), hasData = as.character(), seconds = as.numeric()) %>% data.frame()
# 
# # calculate tiles
# elev_dims <- st_dimensions(elevation_st)
# slope_dims <- st_dimensions(slope_st)
# 
# xmax  <- elev_dims$x$to
# ymax  <- elev_dims$y$to
# xsize <- 10000
# ysize <- 4000
# 
# xnfull      <- round(xmax/xsize, digits = 0) 
# xremainder  <- xmax-(xsize*xnfull)
# ynfull      <- round(ymax/ysize, digits = 0) 
# yremainder  <- ymax-(ysize*ynfull)
# 
# tiles <- data.frame(xoff = rep(seq(1, by = xsize, length.out = xnfull+1), ynfull+1), 
#                     yoff = rep(seq(1, by = ysize, length.out = ynfull+1), each = xnfull+1),
#                     xsize = rep(c(rep(xsize, length.out = xnfull), xremainder), ynfull+1),
#                     ysize = c(rep(ysize, length.out = (xnfull+1)*ynfull), rep(yremainder, xnfull+1)))
# 
# NP_boundaries <- readOGR(dsn = "../data/NP_boundaries/National_Parks__England____Natural_England.shp", 
#                            verbose = TRUE) %>% st_as_sf()
# 
#   site_bound <- NP_boundaries[which(NP_boundaries$NAME == "YORKSHIRE DALES"),] 
#   site_bbox <- st_bbox(site_bound)
#   
#   
#   #define an area that's got a lot of points in it, for testing
#   # site_bbox <- st_bbox(c(xmin = 378152.2, ymin = 470897.0, 
#   #                        xmax = 398152.2, ymax = 490897.0), 
#   #                      crs = st_crs(elevation_st)) #st_bbox(site_bound)
#   site_dims <- st_dimensions(elevation_st[site_bbox])  #get teh dimensions in raster coords of the cropped layer
#   
#   xoff <- site_dims$x$from; yoff <- site_dims$y$from; 
#   xsize <- site_dims$x$to - site_dims$x$from; ysize <- site_dims$y$to - site_dims$y$from
#   # xoff <- tiles$xoff[i]; yoff <- tiles$yoff[i]; xsize <- tiles$xsize[i]; ysize <- tiles$ysize[i]
#   # 
#   rasterio <- list(nXOff = xoff, nYOff = yoff, nXSize = xsize, nYSize = ysize)
```


```{r}
# list of tiles
tilelist <- list.files(path = "../data/tiles_5k", pattern = "*.tif$", ignore.case = T)
length(tilelist)/2

# loop through each tile, test if it has any values, then krige the heck out of it

#for(i in 1:nrow(tiles)) {
  start <- Sys.time()

  
  elev_tile <- read_stars("../data/tiles_5k/elev_eng_5m-449.tif", 
                          proxy = F, quiet = F)
  slope_tile <- read_stars("../data/tiles_5k/slope_eng_5m-449.tif", 
                           proxy = F, quiet = F)
  preds_tile <- c(elev_tile, slope_tile)
  names(preds_tile) <- c("elev", "slope")
  

  
  st_dimensions(preds_tile)
  
  #read to local raster to assess if there are any non-null values
  slope_tile_r <- raster(slope_tile$`slope_eng_5m-449.tif`, crs = st_crs(slope_tile))
  

```


```{r}
  
  if (!is.nan(cellStats(slope_tile_r, mean))) { #test if there are any non-null values
    
    rm(slope_tile_r)  
    
    Gb <- object.size(preds_tile)/1000000000 #calc size in memory
    #print(as.character(c(i = i, xoff = xoff, yoff = xoff, Gb = Gb, hasData = "Data")))
    
    krige.out <- gstat::krige(formula = cbrt_depth ~ elev + slope, 
                              maxdist = 5000, nmax = 10,
                              locations = dat_sf, 
                              newdata = preds_tile, 
                              model = model.geo)
    #backtransform
    krige.out <- krige.out %>% mutate(pred.btf = btf(var1.pred), 
                                      var.btf = btf(var1.var),
                                      LPI = pred.btf - 1.96 * sqrt(var.btf),
                                      UPI = pred.btf + 1.96 * sqrt(var.btf),
                                      interval = UPI - LPI)
    #need to add a line to make tiles directory
    #export
    write_stars(krige.out["var1.pred"], paste0("../outputs/tiles_", parameters$rundate, "/pred_tf_tl", 
                                               i, "_x", xoff, "_y", yoff, ".tif"))
    write_stars(krige.out["var1.var"], paste0("../outputs/tiles_", parameters$rundate, "/var_tl", 
                                              i, "_x", xoff, "_y", yoff, ".tif"))
    write_stars(krige.out["interval"], paste0("../outputs/tiles_", parameters$rundate, "/interval_tl", 
                                              i, "_x", xoff, "_y", yoff, ".tif"))
    
    #record progress and results
    end <- Sys.time(); time <- difftime(end, start, units = "mins")
    tracking <- rbind(tracking, c(i = i, xoff = xoff, yoff = yoff, Gb = Gb, hasData = "Data", minutes = time))
    names(tracking) <- c("i", "xoff", "yoff", "Gb", "hasData", "minutes")
    
    rm(preds_tile, krige.out) # make space in memory
  } else {
    Gb <- object.size(preds_tile)/1000000000 #calc size in memory
    print(as.character(c(i = i, xoff = xoff, yoff = xoff, Gb = Gb, hasData = "noData")))
    end <- Sys.time(); time <- difftime(end, start, units = "mins")
    tracking <- rbind(tracking, c(i = i, xoff = xoff, yoff = yoff, Gb = Gb, isnull = "noData", minutes = NA))
    names(tracking) <- c("i", "xoff", "yoff", "Gb", "hasData", "minutes")
  }
  
}


```
Error in predict.gstat(g, newdata = newdata, block = block, nsim = nsim,  : 
  out of dynamic memory (try local kriging?)
  
  [1] "191"         "100001"      "100001"      "0.640013488"
[using universal kriging]
GDAL Error 5: F:\Users\Bruce\Documents\GitHub\Peat_Depth_Model\england\data\elev_eng_5m.tif: Access window out of range in RasterIO().  Requested (110000,60000) of size 10000x4000 on raster of 114203x128824.Error in CPL_read_gdal(as.character(x), as.character(options), as.character(driver),  : 
  read failure

```{r}
plot(elev_tile)
plot(dat_sf, add = TRUE)

```

```{r}
#crashes r
#plot(dat_sf["depth"], reset = F, border = NA)
# plot(elev_tile, add = T)

# plot(elevation_st[site_bbox])
# plot(st_as_sfc(st_bbox(elevation_st)), add = TRUE, border = "green")

plot(st_as_sfc(st_bbox(elev_tile)), reset = FALSE, border = NA) #this effectively crops the plotting window without the hassle of cropping the data
plot(elev_tile, add = TRUE)
#plot(elevation_st, add = TRUE)
#plot(elevation_st[site_bbox], add = TRUE)
plot(st_as_sfc(st_bbox(elev_tile)), 
     add = TRUE, 
     border = "red")
plot(dat_sf["depth"], add = TRUE, cex = 0.2)
```



```{r}
#check if any of them worked
tilelist <- list.files(path = paste0("../outputs/tiles_", parameters$rundate), full.names = T)
for(i in 1:length(tilelist)){
print(paste(i, tilelist[i]))
try(temp_tile <- read_stars(tilelist[i]))
try(plot(temp_tile))
}
```


```{r}
sf::st_coordinates(x)
sf::st_crs(x)
stars::st_raster_type(x)
```






```{r}
nxoff <- 100000
nyoff <- 50000
rasterio <- list(nXOff = nxoff, nYOff = nyoff, nXSize = 114203-nxoff, nYSize = 128824-nyoff, bands = c(1))
x <- read_stars(parameters$elev_raster_fn, RasterIO = rasterio, proxy = T, quiet = F)
plot(x)
```



```{r}
gstat::krige(formula = cbrt_depth ~ elev + slope, locations = sp_depth, 
                   newdata = preds_st, model = model.geo)
```






```{r}
input.data.gs <- input.data.sp #spdf format

#remove duplicates 
input.data.gs <- remove.duplicates(obj = input.data.gs)

sp_depth <- input.data.gs

```





```{r}
# sp_depth <- sp::spTransform(x = sp_depth, CRSobj = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +a=6377563.396 +rf=299.324975315035 +towgs84=446.448,-125.157,542.06,0.1502,0.247,0.8421,-20.48939999999 +units=m +no_defs")
```


```{r}
# preds_st <- st_transform_proj(x = preds_st, crs = "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +ellps=airy +towgs84=446.448,-125.157,542.06,0.1502,0.247,0.8421,-20.48939999999 +units=m +no_defs")
# 
# preds_st <- st_transform_proj(x = preds_st, crs = proj4string(sp_depth))
```

```{r}
#ggplot() + geom_stars(preds_st) + coord_equal()
```

```{r}
# library(ggplot2)
# ggplot() + geom_stars(data = preds_st) +
#     coord_equal() +
#     facet_wrap(~band) +
#     theme_void() +
#     scale_x_discrete(expand=c(0,0))+
#     scale_y_discrete(expand=c(0,0))
```



## Prediction


```{r}
krige.out <- krige(formula = sqrt_depth ~ elev + slope, 
                   locations = sp_depth,
                   newdata = preds_st,
                   model = model.geo)
```

```{r}
saveRDS(object = krige.out, file = paste0("../outputs/krige.out_", 
                      parameters$area_abbr, "_",
                      rundate, ".rds"))
```


```{r}
class(krige.out)
str(krige.out)
plot(krige.out)
```

```{r}
stars::write_stars()

methods(class = "stars")
```

```{r}
krige.pred.rst <- raster(krige.out$var1.pred)
krige.var.rst <- raster(krige.out$var1.var)

writeRaster(x = krige.pred.rst, 
            filename = paste0("../outputs/krige.pred_", parameters$area_abbr, "_",
                      rundate, ".tif"))

writeRaster(x = krige.var.rst, 
            filename = paste0("../outputs/krige.var_", parameters$area_abbr, "_",
                      rundate, ".tif"))

rm(krige.pred.rst, krige.var.rst)
rm(krige.out)

```







